---
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE,out.extra='trim={0 1cm 0 1cm}') 
```
### **Introduction**
The project looks into the a data set over viewing certain beers and rating them by a criteria.
```{r, message=F, warning=F}
#install packages
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(data.table)
library(corrplot)
library(corrgram)
library(forcats)
```

**Overview of the Data**

The data set goes over beer that were reviewed by people based on characteristic.

The data set has 13 variables/columns as follows:

beer_ABV: alcohol by volume, is the standard measurement, used worldwide, to assess the strength of a particular beer

beer_beerId: Id number that was given for the beer_name

beer_brewerId: Id number that given for the beer_style

beer_name: Name of the beer

beer_style: Style of the beer

review_appearance: How the look of the beer looked to reviewer overall. Scored by 1 out of 5.

review_palette: The range of taste of the beer.Scored by 1 out of 5.

review_overall: The overall review of the beer. Taken the mean of all the other reviews. Scored by 1 out of 5.

review_taste : How the beer tasted. Scored by 1 out of 5.

review_profileName:  The profile name of the reviewer.

review_aroma: The scent of the beer. Scored by 1 out of 5.

review_text: The comment the reviewer left.

review_time: When the reviwer made the review.
```{r}
# puts the data set into variable BP
BP <- read.csv("BeerProject.csv") 
# shows the first 5 rows of the data set
head(BP) 
```

```{r}
# Shows the variable and how many rows the data set has
str(BP)
```

```{r}
# Shows the overall summary of the data set
summary(BP) 
```

### **Cleaning the data set** 

The cleaning process involved checking to see if there were any NA's or blanks in the column. Removing the following columns;review_text, review_time and review_profileName, as I felt those columns would not affect the results of the findings of the this project. Making a function to get the mean review score of the beers as beers were reviewed more then once.
```{r}
colSums(is.na(BP))  # Shows any columns that contains NA's or Blanks
```

```{r}
# functions gets the overall average review score of the beers as beers were reviewed more then once
getmode <- function(v) { 
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v,uniqv)))]
}

beer_ABV_mode<-getmode(BP$beer_ABV)
BP$beer_ABV[which(is.na(BP$beer_ABV))] <- beer_ABV_mode

colSums(is.na(BP))
```

```{r}
# removes columns
BP2 <- BP %>% subset(select=-c(review_text,review_time,review_profileName)) 
```

```{r}
# sees the first 5 rows of the edited data set
head(BP2) 
```

### **Visualizations of the characteristics that were reviewed for the beers.**

The following code chucks shows aggregation and visualizations of characteristics of the beers that were scored. The following visualization were are table that shows the count or the average score of the beers. A point chart or a bar chart to visualize the results from the table. A histogram to shows the frequency of the score, 
```{r}
# counts the different types of beers.Outputs only the first 5 rows
type_df <- BP2 %>% group_by(beer_style) %>% summarize(counts=n()) 
head(type_df)
```

```{r}
# Charts the counts of the type of beers in a bar chart
type_df %>% top_n(n=20) %>% mutate(beer_style = fct_reorder(beer_style,desc(counts))) %>% 
ggplot(aes(x=beer_style,y=counts)) + 
geom_bar(stat = "identity")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```

```{r}
# counts the names of the beers that were reviewed 
name_df <- BP2 %>% group_by(beer_name) %>% summarize(counts=n())
name_df
```

```{r}
# Charts the counts of the name of the beers in a bar chart
name_df %>% top_n(n=20) %>% mutate(beer_name = fct_reorder(beer_name,desc(counts))) %>% 
ggplot(aes(x=beer_name,y=counts)) + 
geom_bar(stat = "identity")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

```{r}
# shows the average alcohol content of the beers. Output only shows the first top 5 of the data set
beer_ABV_df <- BP2 %>% group_by(beer_style) %>%  summarize(mean(beer_ABV)) 
names(beer_ABV_df)[2] <- "Average_Alcohol_content_by_volume"
beer_ABV_df <- beer_ABV_df %>% arrange(desc(Average_Alcohol_content_by_volume))
head(beer_ABV_df)
```

```{r}
# charts the results of the table in a bar chart. Outputs only the top 20
beer_ABV_df %>% top_n(n=20) %>% 
mutate(beer_style = fct_reorder(beer_style, Average_Alcohol_content_by_volume, .desc = TRUE)) %>% ggplot(aes(beer_style,Average_Alcohol_content_by_volume)) +
geom_bar(stat = "identity")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

```{r}
# Shows the overall rating of the beers. Outputs only the first 5 rows of the data set
beer_name_rating <- BP2 %>% group_by(beer_name) %>% summarize(mean(review_overall)) 

names(beer_name_rating)[2] <- "Average_Review_Overall"

beer_name_rating <- beer_name_rating %>% arrange(desc(Average_Review_Overall))
beer_name_rating
```

```{r}
# shows the frequency of each review score 
beer_name_rating %>% ggplot(aes(x=Average_Review_Overall)) + 
geom_histogram(color="black", fill="white",binwidth=1)
```

```{r}
# Shows the average appearance score by beer style. Outputs only the first 5 rows
review_appearance_df <- BP2 %>% group_by(beer_style
) %>%  summarize(mean(review_appearance)) 
names(review_appearance_df)[2] <- "Avg_review_appearance"
review_appearance_df<- review_appearance_df %>% arrange(desc(Avg_review_appearance))
head(review_appearance_df)
```

```{r}
# Charts the results from the table in a point chart. Only output the top 20.
review_appearance_df %>% top_n(n=20)%>%
mutate(beer_style = fct_reorder(beer_style, Avg_review_appearance, .desc = TRUE)) %>% 
ggplot(aes(beer_style,Avg_review_appearance))+geom_point(stat = "identity")+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```


```{r}
# shows the frequency of each review score 
review_appearance_df %>% ggplot(aes(x=Avg_review_appearance)) + 
geom_histogram(color="black", fill="white",binwidth=1)
```

```{r}
# Shows the average review palette by beer style. Outputs only the first 5 rows
review_palette_df <- BP2 %>% group_by(beer_style) %>% summarize(mean(review_palette))
names(review_palette_df)[2] <- "Avg_review_palette"
review_palette_df <- review_palette_df %>% arrange(desc(Avg_review_palette))
head(review_palette_df)
```

```{r}
# Plots the results from the table in a graph. Only outputs the top 20
review_palette_df %>% top_n(n=20) %>% 
mutate(beer_style =  fct_reorder(beer_style, Avg_review_palette, .desc = TRUE)) %>% 
ggplot(aes(beer_style,Avg_review_palette)) + geom_point(stat ="identity" )+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

```{r}
# shows the frequency of each review score
review_palette_df %>% ggplot(aes(x=Avg_review_palette)) + 
geom_histogram(color="black", fill="white",binwidth=1)
```


```{r}
# Shows the average review taste by beer style. Outputs only the first 5 rows
review_taste_df <- BP2 %>% group_by(beer_style) %>% summarize(mean(review_taste))
names(review_taste_df)[2] <- "Avg_review_taste"
review_taste_df <- review_taste_df %>% arrange(desc(Avg_review_taste))
head(review_taste_df)
```

```{r}
# Plots the results from the table in a graph. Only outputs the top 20
review_taste_df %>% top_n(n=20) %>% 
mutate(beer_style =  fct_reorder(beer_style, Avg_review_taste, .desc = TRUE)) %>% 
ggplot(aes(beer_style,Avg_review_taste)) + geom_point(stat ="identity" )+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```


```{r}
# shows the frequency of each review score 
review_taste_df %>% ggplot(aes(x=Avg_review_taste)) + 
geom_histogram(color="black", fill="white",binwidth=1)
```

```{r}
# Shows the average aroma taste by beer style. Outputs only the first 5 rows
review_aroma_df <- BP2 %>% group_by(beer_style) %>% summarize(mean(review_aroma))
names(review_aroma_df)[2] <-"Avg_review_aroma"
review_aroma_df <- review_aroma_df %>% arrange(desc(Avg_review_aroma))
head(review_aroma_df)
```

```{r}
# Plots the results from the table in a graph. Only outputs the top 20
review_aroma_df %>% top_n(n=20) %>% 
mutate(beer_style =  fct_reorder(beer_style,Avg_review_aroma, .desc = TRUE)) %>% 
ggplot(aes(beer_style,Avg_review_aroma)) + geom_point(stat ="identity" )+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

```{r}
# shows the frequency of each review score 
review_aroma_df %>% ggplot(aes(x=Avg_review_aroma)) + 
geom_histogram(color="black", fill="white",binwidth=1)
```


### **Correlation of the variables**
The following code chunks were used to get the correlation of the variables that were reviewed for the beers.
First I made a code that would only get the columns that had the scores regarding the review and put them in a correlation table showing the correlation the variables with each other.  
```{r}
cordata = BP2[,c(6,7,8,9,10)]
corr <- cor(cordata)
corr
```

### **Correlation of the variables**
Here I made correlation plot of the variables to visualize what the table had.
```{r}
corrplot(corr,method = "circle")
```

## **Conclusion**
With the data that has been collected we can see the highest correlation with the overall review was aroma, taste, and palette. A note to take from the analyzing the data set is that NOT all beers were revewived by the same amount which could have affected the correlation given here.